---
title: "Project"
author: "Brian de Korodi"
date: "2024-12-17"
output: pdf_document
---

```{r, message=FALSE, warning=FALSE}
rm(list = ls())
```

### Packages

```{r, message=FALSE, warning=FALSE}
#install.packages("stringr")
#install.packages("dplyr")
#install.packages("httr")
#install.packages("rvest")
#install.packages("ggplot2")
#install.packages("tidyr")
#install.packages("tidyverse")
#install.packages("caret")
#install.packages("randomForest")
library(stringr)
library(dplyr)
library(httr)
library(rvest)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(purrr)
library(caret)
library(randomForest)
```

```{r}
Seasons_folder <- "/Users/bdekorodi/Desktop/Project/Project_data/Seasons"
Playoffs_folder <- "/Users/bdekorodi/Desktop/Project/Project_data/Playoffs"
```

### Extracting data

#### Seasons - Team per 100 possessions

```{r, message=FALSE, warning=FALSE}
Seasons_Team_per100poss <- list()
for (year in 1997:2024) {
  Seasons_file <- file.path(Seasons_folder, paste0("Season_", year, ".html"))
  tryCatch({
    Seasons_Team_per100poss[[paste0("team_per100poss_", year)]] <- read_html(Seasons_file) %>%
      html_nodes(xpath = '//*[@id="per_poss-team"]') %>%
      html_table(header=TRUE) %>%
      .[[1]] %>%
      mutate(Year = year)
  })
}
col_names_Team_per100poss <- c("Team", "Games_Played", "Minutes_Played", "FGMr", "FGAr", "FG%","3PMr", "3PAr", "3P%", "2PMr", "2PAr", "2P%", "FTMr", "FTAr", "FT%", "ORBr", "DRBr", "TRBr","ASTr", "STLr", "BLKr", "TOVr", "PFr", "ORtg", "Year")
Seasons_Team_per100poss <- lapply(Seasons_Team_per100poss, function(df) {
  df <- df[, -1, drop = FALSE]
  colnames(df) <- col_names_Team_per100poss[seq_len(ncol(df))]
    if ("Team" %in% colnames(df)) {
    df$Team <- str_remove(df$Team, "\\*$")
    df$Team <- str_replace_all(df$Team, c(
      "New Jersey Nets" = "Brooklyn Nets",
      "Seattle SuperSonics" = "Oklahoma City Thunder",
      "Vancouver Grizzlies" = "Memphis Grizzlies",
      "Washington Bullets" = "Washington Wizards",
      "New Orleans Hornets" = "New Orleans Pelicans",
      "New Orleans/Oklahoma City Hornets" = "New Orleans Pelicans",
      "Charlotte Bobcats" = "Charlotte Hornets"
    ))
    df <- df %>%
    mutate(
      Team = as.character(Team),
      Year = as.integer(Year),
      across(.cols = -c(Team, Year), .fns = as.numeric)
      )
  }
  return(df)
  
})
team_per100poss_rm_row_21 <- c("team_per100poss_2005", "team_per100poss_2006", "team_per100poss_2007", "team_per100poss_2008", "team_per100poss_2009", "team_per100poss_2010", "team_per100poss_2011", "team_per100poss_2012", "team_per100poss_2013", "team_per100poss_2014", "team_per100poss_2015", "team_per100poss_2016", "team_per100poss_2017", "team_per100poss_2018", "team_per100poss_2019", "team_per100poss_2020", "team_per100poss_2021", "team_per100poss_2022", "team_per100poss_2023", "team_per100poss_2024")

for (name in team_per100poss_rm_row_21) {
  Seasons_Team_per100poss[[name]] <- Seasons_Team_per100poss[[name]][-c(21), ]
}
```

#### Seasons - Opponent per 100 possessions

```{r, message=FALSE, warning=FALSE}
Seasons_Opp_per100poss <- list()
for (year in 1997:2024) {
  Seasons_file <- file.path(Seasons_folder, paste0("Season_", year, ".html"))
  tryCatch({
    Seasons_Opp_per100poss[[paste0("opp_per100poss_", year)]] <- read_html(Seasons_file) %>%
      html_nodes(xpath = '//*[@id="per_poss-opponent"]') %>%
      html_table(header=TRUE) %>%
      .[[1]] %>%
      mutate(Year = year)
  })
}
col_names_Opp_per100poss <- c("Team", "Games_Played", "Minutes_Played", "oppFGMr", "oppFGAr", "oppFG%","opp3PMr", "opp3PAr", "opp3P%", "opp2PMr", "opp2PAr", "opp2P%", "oppFTMr", "oppFTAr", "oppFT%", "oppORBr", "oppDRBr", "oppTRBr","oppASTr", "oppSTLr", "oppBLKr", "oppTOVr", "oppPFr", "DRtg", "Year")
Seasons_Opp_per100poss <- lapply(Seasons_Opp_per100poss, function(df) {
  df <- df[, -1, drop = FALSE]
  colnames(df) <- col_names_Opp_per100poss[seq_len(ncol(df))]
    if ("Team" %in% colnames(df)) {
    df$Team <- str_remove(df$Team, "\\*$")
    df$Team <- str_replace_all(df$Team, c(
      "New Jersey Nets" = "Brooklyn Nets",
      "Seattle SuperSonics" = "Oklahoma City Thunder",
      "Vancouver Grizzlies" = "Memphis Grizzlies",
      "Washington Bullets" = "Washington Wizards",
      "New Orleans Hornets" = "New Orleans Pelicans",
      "New Orleans/Oklahoma City Hornets" = "New Orleans Pelicans",
      "Charlotte Bobcats" = "Charlotte Hornets"
    ))
    df <- df %>%
    mutate(
      Team = as.character(Team),
      Year = as.integer(Year),
      across(.cols = -c(Team, Year), .fns = as.numeric)
      )
  }
  return(df)
})
```

#### Seasons - Team advanced

```{r, message=FALSE, warning=FALSE}
Seasons_Team_advanced <- list()

for (year in 1997:2024) {
  Seasons_file <- file.path(Seasons_folder, paste0("Season_", year, ".html"))
  
  tryCatch({
    Seasons_Team_advanced[[paste0("team_advanced_", year)]] <- read_html(Seasons_file) %>%
      html_nodes(xpath = '//*[@id="advanced-team"]') %>%
      html_table(fill = TRUE) %>%
      .[[1]]
    col_names <- Seasons_Team_advanced[[paste0("team_advanced_", year)]][1, ]
    Seasons_Team_advanced[[paste0("team_advanced_", year)]] <- Seasons_Team_advanced[[paste0("team_advanced_", year)]][-1, ]
    col_names <- ifelse(is.na(col_names) | col_names == "", 
                        paste0("Col", seq_along(col_names)), 
                        col_names)
    colnames(Seasons_Team_advanced[[paste0("team_advanced_", year)]]) <- col_names
  })
}

col_names_Team_advanced <- c("Rank", "Team", "Average_age", "Wins", "Losses", "Expected_wins", "Expected_Losses", "MOV", "SOS", "SRS", "ORtg", "DRtg", "NRtg", "Pace", "FTA_per_FGA", "%3PA", "TS%", "Col18", "eFG%", "TOV%","ORB%", "FT_per_FGA", "Col23", "oppeFG%", "oppTOV%", "DRB%", "oppFT_per_FGA", "Col28", "Arena", "Attendance", "Attendance_per_Game")
Seasons_Team_advanced <- lapply(Seasons_Team_advanced, function(df) {
  colnames(df) <- col_names_Team_advanced[seq_len(ncol(df))]
    if ("Team" %in% colnames(df)) {
    df$Team <- str_remove(df$Team, "\\*$")
    df$Team <- str_replace_all(df$Team, c(
      "New Jersey Nets" = "Brooklyn Nets",
      "Seattle SuperSonics" = "Oklahoma City Thunder",
      "Vancouver Grizzlies" = "Memphis Grizzlies",
      "Washington Bullets" = "Washington Wizards",
      "New Orleans Hornets" = "New Orleans Pelicans",
      "New Orleans/Oklahoma City Hornets" = "New Orleans Pelicans",
      "Charlotte Bobcats" = "Charlotte Hornets"
    ))
    }
  df <- df %>%
      select(-c(Col18, Col23, Col28, Rank))
    df <- df %>%
  mutate(
    Attendance = as.numeric(gsub(",", "", Attendance)),
    Attendance_per_Game = as.numeric(gsub(",", "", Attendance_per_Game))
  )
  df <- df %>%
    mutate(
      Team = as.character(Team),
      Arena = as.character(Arena),
      across(.cols = -c(Team, Arena), .fns = as.numeric)
      )
  return(df)
})

team_advanced_rm_row_16_17 <- c("team_advanced_2005", "team_advanced_2006", "team_advanced_2007", "team_advanced_2008", "team_advanced_2009", "team_advanced_2010", "team_advanced_2011", "team_advanced_2012", "team_advanced_2013", "team_advanced_2014", "team_advanced_2015", "team_advanced_2016", "team_advanced_2017", "team_advanced_2018", "team_advanced_2019", "team_advanced_2020", "team_advanced_2021", "team_advanced_2022", "team_advanced_2023", "team_advanced_2024")

for (name in team_advanced_rm_row_16_17) {
  Seasons_Team_advanced[[name]] <- Seasons_Team_advanced[[name]][-c(16, 17), ]
}
```

#### Seasons - Team shooting

Basketball reference added Layup %FGA and total made layups to shooting data starting the 01-02 season.

```{r, message=FALSE, warning=FALSE}

Seasons_Team_shooting_pre_2002 <- list()
Seasons_Team_shooting_post_2002 <- list()

col_names_Team_shooting_pre_2002 <- c("Rank", "Team", "Games_Played", "Minutes_Played", "FG%", "Average_FGA_distance", "Col7", "2P%FGA", "0_3%FGA", "3_10%FGA", "10_16%FGA", "16_3P%FGA", "3P%FGA", "Col14", "2P%", "0_3FG%", "3_10FG%", "10_16FG%","16_3PFG%", "3P%", "Col21", "2P%FGAst", "3P%FGAst", "Col24", "Dunks%FGA", "Made_dunks", "Col27", "Corner%3PA", "Corner3P%", "Col30", "Attempted_heaves", "Made_heaves")

col_names_Team_shooting_post_2002 <- c("Rank", "Team", "Games_Played", "Minutes_Played", "FG%", "Average_FGA_distance", "Col7", "2P%FGA", "0_3%FGA", "3_10%FGA", "10_16%FGA", "16_3P%FGA", "3P%FGA", "Col14", "2P%", "0_3FG%", "3_10FG%", "10_16FG%", "16_3PFG%", "3P%", "Col21", "2P%FGAst", "3P%FGAst", "Col24", "Dunks%FGA", "Made_dunks", "Col27", "Layups%FGA", "Made_Layups", "Col33", "Corner%3PA", "Corner3P%", "Col30", "Attempted_heaves", "Made_heaves")

for (year in 1997:2001) {
  Seasons_file <- file.path(Seasons_folder, paste0("Season_", year, ".html"))
  
  tryCatch({
    df <- read_html(Seasons_file) %>%
      html_nodes(xpath = '//*[@id="shooting-team"]') %>%
      html_table(fill = TRUE) %>%
      .[[1]]
    col_names <- df[1, ]
    df <- df[-1, ]
    col_names <- ifelse(is.na(col_names) | col_names == "", 
                        paste0("Col", seq_along(col_names)), 
                        col_names)
    colnames(df) <- col_names
    colnames(df) <- col_names_Team_shooting_pre_2002[seq_len(ncol(df))]
    Seasons_Team_shooting_pre_2002[[paste0("team_shooting_", year)]] <- df
  }, error = function(e) {
    message(paste("Error processing year:", year))
  })
}

for (year in 2002:2024) {
  Seasons_file <- file.path(Seasons_folder, paste0("Season_", year, ".html"))
  
  tryCatch({
    df <- read_html(Seasons_file) %>%
      html_nodes(xpath = '//*[@id="shooting-team"]') %>%
      html_table(fill = TRUE) %>%
      .[[1]]
    col_names <- df[1, ]
    df <- df[-1, ]
    col_names <- ifelse(is.na(col_names) | col_names == "", 
                        paste0("Col", seq_along(col_names)), 
                        col_names)
    colnames(df) <- col_names
    colnames(df) <- col_names_Team_shooting_post_2002[seq_len(ncol(df))]
    df <- df[, !colnames(df) %in% c("Layups%FGA", "Made_Layups", "Col33")]
    Seasons_Team_shooting_post_2002[[paste0("team_shooting_", year)]] <- df
  })
}

Seasons_Team_shooting <- c(Seasons_Team_shooting_pre_2002, Seasons_Team_shooting_post_2002)

names(Seasons_Team_shooting) <- paste0("team_shooting_", 1997:2024)

Seasons_Team_shooting <- lapply(Seasons_Team_shooting, function(df) {
    df$Team <- str_remove(df$Team, "\\*$")
    df$Team <- str_replace_all(df$Team, c(
      "New Jersey Nets" = "Brooklyn Nets",
      "Seattle SuperSonics" = "Oklahoma City Thunder",
      "Vancouver Grizzlies" = "Memphis Grizzlies",
      "Washington Bullets" = "Washington Wizards",
      "New Orleans Hornets" = "New Orleans Pelicans",
      "New Orleans/Oklahoma City Hornets" = "New Orleans Pelicans",
      "Charlotte Bobcats" = "Charlotte Hornets"
    ))
    df <- df[, !colnames(df) %in% c("Rank", "Col7", "Col14", "Col21", "Col24", "Col27", "Col30")]
    df <- df %>%
    mutate(
      Team = as.character(Team),
      across(.cols = -Team, .fns = as.numeric)
    )
  return(df)
})
team_shooting_rm_row_16_17 <- c("team_shooting_2005", "team_shooting_2006", "team_shooting_2007", "team_shooting_2008", "team_shooting_2009", "team_shooting_2010", "team_shooting_2011", "team_shooting_2012", "team_shooting_2013", "team_shooting_2014", "team_shooting_2015", "team_shooting_2016", "team_shooting_2017", "team_shooting_2018", "team_shooting_2019", "team_shooting_2020", "team_shooting_2021", "team_shooting_2022", "team_shooting_2023", "team_shooting_2024")

for (name in team_shooting_rm_row_16_17) {
  Seasons_Team_shooting[[name]] <- Seasons_Team_shooting[[name]][-c(16, 17), ]
}

rm(Seasons_Team_shooting_post_2002, Seasons_Team_shooting_pre_2002, df, col_names)
```

#### Seasons - Opponent shooting

```{r, message=FALSE, warning=FALSE}

Seasons_Opp_shooting_pre_2002 <- list()
Seasons_Opp_shooting_post_2002 <- list()

col_names_Opp_shooting_pre_2002 <- c("Rank", "Team", "Games_Played", "Minutes_Played", "oppFG%", "oppAverage_FGA_distance", "Col7", "opp2P%FGA", "opp0_3%FGA", "opp3_10%FGA", "opp10_16%FGA", "opp16_3P%FGA", "opp3P%FGA", "Col14", "opp2P%","opp0_3FG%", "opp3_10FG%", "opp10_16FG%", "opp16_3PFG%", "opp3P%", "Col21", "opp2P%FGAst","opp3P%FGAst", "Col24", "oppDunks%FGA", "oppMade_dunks", "Col27", "oppCorner%3PA", "oppCorner3P%")

col_names_Opp_shooting_post_2002 <- c("Rank", "Team", "Games_Played", "Minutes_Played", "oppFG%", "oppAverage_FGA_distance", "Col7", "opp2P%FGA", "opp0_3%FGA", "opp3_10%FGA", "opp10_16%FGA", "opp16_3P%FGA", "opp3P%FGA", "Col14", "opp2P%","opp0_3FG%", "opp3_10FG%", "opp10_16FG%", "opp16_3PFG%", "opp3P%", "Col21", "opp2P%FGAst","opp3P%FGAst", "Col24", "oppDunks%FGA", "oppMade_dunks", "Col27", "oppLayups%FGA", "oppMade_Layups", "Col30", "oppCorner%3PA", "oppCorner3P%")

for (year in 1997:2001) {
  Seasons_file <- file.path(Seasons_folder, paste0("Season_", year, ".html"))
  
  tryCatch({
    df <- read_html(Seasons_file) %>%
      html_nodes(xpath = '//*[@id="shooting-opponent"]') %>%
      html_table(fill = TRUE) %>%
      .[[1]]
    col_names <- df[1, ]
    df <- df[-1, ]
    col_names <- ifelse(is.na(col_names) | col_names == "", 
                        paste0("Col", seq_along(col_names)), 
                        col_names)
    colnames(df) <- col_names
    colnames(df) <- col_names_Opp_shooting_pre_2002[seq_len(ncol(df))]
    Seasons_Opp_shooting_pre_2002[[paste0("opp_shooting_", year)]] <- df
  })
}

for (year in 2002:2024) {
  Seasons_file <- file.path(Seasons_folder, paste0("Season_", year, ".html"))
  
  tryCatch({
    df <- read_html(Seasons_file) %>%
      html_nodes(xpath = '//*[@id="shooting-opponent"]') %>%
      html_table(fill = TRUE) %>%
      .[[1]]
    col_names <- df[1, ]
    df <- df[-1, ]
    col_names <- ifelse(is.na(col_names) | col_names == "", 
                        paste0("Col", seq_along(col_names)), 
                        col_names)
    colnames(df) <- col_names
    colnames(df) <- col_names_Opp_shooting_post_2002[seq_len(ncol(df))]
    df <- df[, !colnames(df) %in% c("oppLayups%FGA", "oppMade_Layups", "Col30", "Col33")]
    Seasons_Opp_shooting_post_2002[[paste0("opp_shooting_", year)]] <- df
  })
}

Seasons_Opp_shooting <- c(Seasons_Opp_shooting_pre_2002, Seasons_Opp_shooting_post_2002)

names(Seasons_Opp_shooting) <- paste0("opp_shooting_", 1997:2024)

Seasons_Opp_shooting <- lapply(Seasons_Opp_shooting, function(df) {
    df$Team <- str_remove(df$Team, "\\*$")
    df$Team <- str_replace_all(df$Team, c(
      "New Jersey Nets" = "Brooklyn Nets",
      "Seattle SuperSonics" = "Oklahoma City Thunder",
      "Vancouver Grizzlies" = "Memphis Grizzlies",
      "Washington Bullets" = "Washington Wizards",
      "New Orleans Hornets" = "New Orleans Pelicans",
      "New Orleans/Oklahoma City Hornets" = "New Orleans Pelicans",
      "Charlotte Bobcats" = "Charlotte Hornets"
    ))
    df <- df[, !colnames(df) %in% c("Rank", "Col7", "Col14", "Col21", "Col24", "Col27")]
    df <- df %>%
    mutate(
      Team = as.character(Team),
      across(.cols = -Team, .fns = as.numeric)
    )
return(df)
})

rm(Seasons_Opp_shooting_post_2002, Seasons_Opp_shooting_pre_2002, df, col_names)
```

#### Joining the datasets from each season by team name

```{r, message=FALSE, warning=FALSE}
clean_column_names <- function(lst) {
  lapply(lst, function(df) {
    colnames(df) <- make.names(colnames(df), unique = TRUE)
    return(df)
  })
}
Seasons_Team_per100poss <- clean_column_names(Seasons_Team_per100poss)
Seasons_Opp_per100poss <- clean_column_names(Seasons_Opp_per100poss)
Seasons_Team_advanced <- clean_column_names(Seasons_Team_advanced)
Seasons_Team_shooting <- clean_column_names(Seasons_Team_shooting)
Seasons_Opp_shooting <- clean_column_names(Seasons_Opp_shooting)

all_team_names <- unique(unlist(lapply(Seasons_Team_per100poss, function(df) df$Team)))

Seasons <- list()

for (year in 1997:2024) {
  Seasons[[paste0("Season_", year)]] <- Seasons_Team_per100poss[[paste0("team_per100poss_", year)]] %>%
    full_join(Seasons_Opp_per100poss[[paste0("opp_per100poss_", year)]], by = "Team") %>%
    full_join(Seasons_Team_advanced[[paste0("team_advanced_", year)]], by = "Team") %>%
    full_join(Seasons_Team_shooting[[paste0("team_shooting_", year)]], by = "Team") %>%
    full_join(Seasons_Opp_shooting[[paste0("opp_shooting_", year)]], by = "Team")%>%
    rename_with(~ gsub("X", "", .)) %>%
    rename("Games_Played" = "Games_Played.x",
           "Minutes_Played" = "Minutes_Played.x",
           "FG%" = "FG..x",
           "2P%" = "2P..x",
           "3P%" = "3P..x",
           "FT%" = "FT.",
           "ORtg" = "ORtg.x",
           "Year" = "Year.x",
           "oppFG%" = "oppFG..x",
           "opp2P%" = "opp2P..x",
           "opp3P%" = "opp3P..x",
           "oppFT%" = "oppFT.",
           "DRB%" = "DRB.",
           "DRtg" = "DRtg.x",
           "%3PA" = ".3PA",
           "TS%" = "TS.",
           "eFG%" = "eFG.",
           "TOV%" = "TOV.",
           "ORB%" = "ORB.",
           "oppeFG%" = "oppeFG.",
           "oppTOV%" = "oppTOV.",
           "2P%FGA" = "2P.FGA",
           "0_3%FGA" = "0_3.FGA",
           "3_10%FGA" = "3_10.FGA",
           "10_16%FGA" = "10_16.FGA",
           "16_3P%FGA" = "16_3P.FGA",
           "3P%FGA" = "3P.FGA",
           "0_3FG%" = "0_3FG.",
           "3_10FG%" = "3_10FG.",
           "10_16FG%" = "10_16FG.",
           "16_3PFG%" = "16_3PFG.",
           "2P%FGAst" = "2P.FGAst",
           "3P%FGAst" = "3P.FGAst",
           "Dunks%FGA" = "Dunks.FGA",
           "Corner%3PA" = "Corner.3PA",
           "Corner3P%" = "Corner3P.",
           "opp2P%FGA" = "opp2P.FGA",
           "opp0_3%FGA" = "opp0_3.FGA",
           "opp3_10%FGA" = "opp3_10.FGA",
           "opp10_16%FGA" = "opp10_16.FGA",
           "opp16_3P%FGA" = "opp16_3P.FGA",
           "opp3P%FGA" = "opp3P.FGA",
           "opp0_3FG%" = "opp0_3FG.",
           "opp3_10FG%" = "opp3_10FG.",
           "opp10_16FG%" = "opp10_16FG.",
           "opp16_3PFG%" = "opp16_3PFG.",
           "opp2P%FGAst" = "opp2P.FGAst",
           "opp3P%FGAst" = "opp3P.FGAst",
           "oppDunks%FGA" = "oppDunks.FGA",
           "oppCorner%3PA" = "oppCorner.3PA",
           "oppCorner3P%" = "oppCorner3P.") %>%
    select(-contains("Games_Played.y"), 
           -contains("Minutes_Played.y"),
           -contains("Year.y"), 
           -contains("Rank.x"),
           -contains("ORtg.y"), 
           -contains("DRtg.y"),
           -contains("Games_Played.x.x"), 
           -contains("Minutes_Played.x.x"),
           -contains("FG..y"), 
           -contains("2P..y"), 
           -contains("3P..y"),
           -contains("Games_Played.y.y"), 
           -contains("Minutes_Played.y.y"), 
           -contains("oppFG..y")) %>%
    mutate(Attendance = gsub(",", "", Attendance),
           Attendance_per_Game = gsub(",", "", Attendance_per_Game)) %>%
    mutate(across(.cols = -c(Arena, Team), .fns = as.numeric)) %>%
    right_join(data.frame(Team = all_team_names), by = "Team")
}
```

#### Playoffs Series - Four Factors

```{r, message=FALSE, warning=FALSE}
Playoffs_FourFactors <- list()
Playoff_series <- c("Finals", "ECF", "WCF", 
                    "ECR2_1", "ECR2_2", "WCR2_1", "WCR2_2", 
                    "ECR1_1", "ECR1_2", "ECR1_3", "ECR1_4", 
                    "WCR1_1", "WCR1_2", "WCR1_3", "WCR1_4")

for (year in 1997:2024) {
  Playoffs_year_folder <- file.path(Playoffs_folder, as.character(year))
  
  for (file in Playoff_series) {
    Playoffs_file <- file.path(Playoffs_year_folder, paste0(file, "_", year, ".html"))
    
    tryCatch({
      raw_html <- readLines(Playoffs_file, warn = FALSE) %>%
        paste(collapse = "\n") %>%
        str_remove_all("<!--|-->")
      Playoffs_FourFactors[[paste0(file, "_", year)]] <- read_html(raw_html) %>%
        html_nodes(xpath = '//*[@id="four_factors"]') %>%
        html_table(fill = TRUE) %>%
        .[[1]]
    }, error = function(e) {})
  }
}

team_name_mapping <- c(
  "CHI" = "Chicago Bulls",
  "BOS" = "Boston Celtics",
  "LAL" = "Los Angeles Lakers",
  "MIA" = "Miami Heat",
  "GSW" = "Golden State Warriors",
  "CLE" = "Cleveland Cavaliers",
  "SA" = "San Antonio Spurs",
  "TOR" = "Toronto Raptors",
  "NYK" = "New York Knicks",
  "OKC" = "Oklahoma City Thunder",
  "MIL" = "Milwaukee Bucks",
  "HOU" = "Houston Rockets",
  "PHI" = "Philadelphia 76ers",
  "SAS" = "San Antonio Spurs",
  "MIN" = "Minnesota Timberwolves",
  "DET" = "Detroit Pistons",
  "IND" = "Indiana Pacers",
  "PHX" = "Phoenix Suns",
  "PHO" = "Phoenix Suns",
  "ORL" = "Orlando Magic",
  "DAL" = "Dallas Mavericks",
  "POR" = "Portland Trail Blazers",
  "SAC" = "Sacramento Kings",
  "WAS" = "Washington Wizards",
  "DEN" = "Denver Nuggets",
  "UTA" = "Utah Jazz",
  "CHA" = "Charlotte Hornets",
  "CHH" = "Charlotte Hornets",
  "CHO" = "Charlotte Hornets",
  "ATL" = "Atlanta Hawks",
  "LAC" = "Los Angeles Clippers",
  "MEM" = "Memphis Grizzlies",
  "NJN" = "Brooklyn Nets",
  "BRK" = "Brooklyn Nets",
  "SEA" = "Oklahoma City Thunder",
  "VAN" = "Memphis Grizzlies",
  "WSB" = "Washington Wizards",
  "NOH" = "New Orleans Pelicans",
  "NOP" = "New Orleans Pelicans")

Playoffs_FourFactors <- lapply(names(Playoffs_FourFactors), function(name) {
  df <- Playoffs_FourFactors[[name]]
      colnames(df)[1] <- "Result"
      colnames(df)[2] <- "Series_Pace"
      colnames(df)[3] <- "Series_eFGpct"
      colnames(df)[4] <- "Series_TOVpct"
      colnames(df)[5] <- "Series_ORBpct"
      colnames(df)[6] <- "Series_FTperFGA"
      colnames(df)[7] <- "Series_ORtg"
      colnames(df)[8] <- "Series_PTSperG"
      df <- df[-1, ]
      df$Series_Pace <- as.numeric(df$Series_Pace)
      df$Series_eFGpct <- as.numeric(df$Series_eFGpct)
      df$`Series_TOVpct` <- as.numeric(df$Series_TOVpct)
      df$`Series_ORBpct` <- as.numeric(df$Series_ORBpct)
      df$Series_FTperFGA <- as.numeric(df$Series_FTperFGA)
      df$Series_ORtg <- as.numeric(df$Series_ORtg)
      df$Series_PTSperG <- as.numeric(df$Series_PTSperG)
  df$Year <- as.integer(str_extract(name, "\\d{4}"))
  df$Series <- str_extract(name, "^[^_]+")
  df$Team <- str_sub(df$Result, 1, 3)
  df$Team <- sapply(df$Team, function(team) team_name_mapping[team])
  df$`Series_Victory_Margin` <- str_extract(df$Result, "\\((\\d+)-(\\d+)\\)") %>%
      str_replace_all("[()]", "") %>%
      str_split("-") %>%
      sapply(function(x) as.integer(x[1]) - as.integer(x[2]))
  df$`Series_Net_Rating` <- c(df$Series_ORtg[1] - df$Series_ORtg[2], df$Series_ORtg[2] -                                   df$Series_ORtg[1])
  df$`Series_Win/Loss` <- c(1, 0)
  df$`Series_Abs_Net_Rating` <- abs(df$`Series_Net_Rating`)
  return(df)
  }) %>% 
setNames(names(Playoffs_FourFactors))
```

```{r, message=FALSE, warning=FALSE}
rm(col_names_Opp_per100poss, col_names_Opp_shooting_post_2002, col_names_Opp_shooting_pre_2002, col_names_Team_advanced, col_names_Team_per100poss, col_names_Team_shooting_post_2002, col_names_Team_shooting_pre_2002, name, team_advanced_rm_row_16_17, team_per100poss_rm_row_21, team_shooting_rm_row_16_17, year, team_name_mapping, file, all_team_names, Playoffs_file, Playoffs_folder, Playoffs_year_folder, raw_html, Seasons_file, Seasons_folder)
```

#### Joining regular season data to playoff series data

```{r, message=FALSE, warning=FALSE}
Playoffs <- list()

for (year in 1997:2024) {
  for (file in Playoff_series) {
    
    playoff_teams <- unique(Playoffs_FourFactors[[paste0(file, "_", year)]][["Team"]])
    
    season_data_filtered <- Seasons[[paste0("Season_", year)]] %>%
      filter(Team %in% playoff_teams)
    
    Playoffs[[paste0(file, "_", year)]] <- left_join(
      Playoffs_FourFactors[[paste0(file, "_", year)]],
      season_data_filtered,
      by = c("Team", "Year")
    )
  }
}
```

```{r, message=FALSE, warning=FALSE}
rm(col_names_Opp_per100poss, col_names_Opp_shooting_post_2002, col_names_Opp_shooting_pre_2002, col_names_Team_advanced, col_names_Team_per100poss, col_names_Team_shooting_post_2002, col_names_Team_shooting_pre_2002, name, team_advanced_rm_row_16_17, team_per100poss_rm_row_21, team_shooting_rm_row_16_17, year, team_name_mapping, file, all_team_names, Playoffs_file, Playoffs_folder, Playoffs_year_folder, raw_html, Seasons_file, Seasons_folder, playoff_teams, season_data_filtered)
```

### Analysis

Our aim is to predict

1)  Who will likely win a playoff matchup

2)  In how many games they will clinch the series

We use playoff series net rating between two teams as an indicator to predict the likelihood of winning the series and in how many games.

So how do we predict the net rating of a playoff series?

-Difference in overall team quality (regular season Net Rating or SRS). This will obviously be a strong predictor of playoff series outcomes. It can also be used as a control variable to isolate the effect of matchup-related predictors

-How often does the team with the best player in the series win? How important is a better supporting cast of role players? This will have to be part of an extension of the current study due to time constraints.

-Playtype factors. Also will have to be part of an extension of the current study due to time constraints.

-Clutch performance factors, particularly as a predictor of the winner of close series (close net ratings). Also will have to be part of an extension of the current study due to time constraints.

-Matchup principal areas indicators. For example we might try to generate indicators of how the opponent's rim protection matches up against the offense. Comparing a defense's foul rate or block rate with an offense's rim FGA rate for example. Or even block rate with 3PA rate (a team who blocks more shots might be more focused on rim protection and thus have their bigs in the paint more, which might create an advantage for an opponent who shoots many threes or many midrangers

Here we will use regular season team data from basketball reference, with a particular emphasis on shooting data and Four Factors.

So the method is, for each playoff series, to go fetch the regular season data for the corresponding teams, and to make transformations on the regular season data so that we get the differentials between the teams' stats.

First, we create differential that measure the difference between Team A's offensive/defensive performance against the league and Team B's defensive/offensive performance against the league, and vice versa.

The weighted shooting differentials measure, for each shot location zone, 1) the differential in efficiency weighted by the proportion of the offense's shots that come from that zone (basically how important that zone is for the offense), 2) the differential in efficiency weighted by the proportion of the defense's shots defended that come from that zone (basically how important that zone is for the defense

```{r}
team_stats <- c("ORtg", "FGMr", "FGAr", "FG%", "3PMr", "3PAr", "3P%", "2PMr", "2PAr", "2P%", "FTAr", "ORBr", "TRBr", "ASTr", "oppSTLr", "oppBLKr", "TOVr", "oppPFr", "Average_age", "Pace", "eFG%", "TOV%", "ORB%", "FT_per_FGA", "Attendance_per_game", "Average_FGA_distance", "2P%FGA", "0_3%FGA", "3_10%FGA", "10_16%FGA", "16_3P%FGA", "3P%FGA", "0_3FG%", "3_10FG%", "10_16FG%", "16_3PFG%", "2P%FGAst", "3P%FGAst", "Dunks%FGA", "Made_dunks", "Corner%3PA", "Corner3P%")
opponent_stats <- c("DRtg", "oppFGMr", "oppFGAr", "oppFG%", "opp3PMr", "opp3PAr", "opp3P%", "opp2PMr", "opp2PAr", "opp2P%", "oppFTAr", "oppORBr", "oppTRBr", "oppASTr", "STLr", "BLKr", "oppTOVr", "PFr", "Average_age", "Pace", "oppeFG%", "oppTOV%", "DRB%", "oppFT_per_FGA", "Attendance_per_game", "oppAverage_FGA_distance", "opp2P%FGA", "opp0_3%FGA", "opp3_10%FGA", "opp10_16%FGA", "opp16_3P%FGA", "opp3P%FGA", "opp0_3FG%", "opp3_10FG%", "opp10_16FG%", "opp16_3PFG%", "opp2P%FGAst", "opp3P%FGAst", "oppDunks%FGA", "oppMade_dunks", "oppCorner%3PA", "oppCorner3P%")

for (name in names(Playoffs)) {
  if (!is.null(Playoffs[[name]]) && nrow(Playoffs[[name]]) == 2) {
    
    df <- Playoffs[[name]]
    
    for (i in seq_along(team_stats)) {
      team_col <- team_stats[i]
      opp_col <- opponent_stats[i]
      new_col <- paste0("Diff_", team_col)
      
      if (team_col %in% names(df) && opp_col %in% names(df)) {

        df[[new_col]] <- c(
          df[[team_col]][1] - df[[opp_col]][2], 
          df[[team_col]][2] - df[[opp_col]][1]   
        )
      }
    }
    Playoffs[[name]] <- df
  }
}
```

Now that we have generated metrics that capture difference in overall team quality, difference in fundamentals, difference in shooting efficiency, and weighted differences in shooting efficiency, we want to create a learning model with Series Net Rating or Series Offensive Rating as the target variable.

Careful, we want to control for overall team quality first. We might want to find out which team quality indicator best correlates historically with Series Net rating, and divide Series Net rating by it.

```{r}
Playoffs_combined <- do.call(rbind, Playoffs)
```

```{r}
# Plot 1: Series eFG% vs eFG%
ggplot(Playoffs_combined, aes(x = `eFG%`, y = Series_eFGpct)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  ggtitle("Series eFG% vs eFG%") +
  xlab("Regular Season eFG%") +
  ylab("Series eFG%")

# Plot 2: Series TOV% vs TOV%
ggplot(Playoffs_combined, aes(x = `TOV%`, y = Series_TOVpct)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  ggtitle("Series TOV% vs TOV%") +
  xlab("Regular Season TOV%") +
  ylab("Series TOV%")

# Plot 3: Series ORB% vs ORB%
ggplot(Playoffs_combined, aes(x = `ORB%`, y = Series_ORBpct)) +
  geom_point() +
  geom_smooth(method = "lm", color = "green") +
  ggtitle("Series ORB% vs ORB%") +
  xlab("Regular Season ORB%") +
  ylab("Series ORB%")

# Plot 4: Series FT per FGA vs FT per FGA
ggplot(Playoffs_combined, aes(x = `FT_per_FGA`, y = Series_FTperFGA)) +
  geom_point() +
  geom_smooth(method = "lm", color = "purple") +
  ggtitle("Series FT/FGA vs FT/FGA") +
  xlab("Regular Season FT per FGA") +
  ylab("Series FT per FGA")

```

```{r}
# Linear regression model
model <- lm(ORtg ~ `eFG%` + `TOV%` + `ORB%` + FT_per_FGA, data = Playoffs_combined)

# Display the summary of the model
summary(model)
```

```{r}
# Linear regression model
model <- lm(Series_ORtg ~ `Series_eFGpct` + `Series_TOVpct` + `Series_ORBpct` + Series_FTperFGA, data = Playoffs_combined)

# Display the summary of the model
summary(model)
```

```{r}
# Linear regression model
model <- lm(Series_ORtg ~ `eFG%` + `TOV%` + `ORB%` + FT_per_FGA, data = Playoffs_combined)

# Display the summary of the model
summary(model)

```

Better offensive teams tend to perform better offensively in the playoffs. However overall quality explains much less of playoff results. Firstly the free throw effect. But we also get a signal that there must be some matchup-related explanation.

```{r}

Playoffs_combined <- Playoffs_combined %>%
  select(-c(Result, Series_Pace, Series_eFGpct, Series_ORBpct, Series_TOVpct, Series_FTperFGA, Series_Abs_Net_Rating, Series_PTSperG, Series_Victory_Margin, Series_Net_Rating, `Series_Win/Loss`, Series_Abs_Net_Rating, Team, Wins, Losses, Expected_wins, Expected_Losses, MOV, SOS, SRS, NRtg, ORtg))

Playoffs_combined <- na.omit(Playoffs_combined)

# Set a seed for reproducibility
set.seed(42)

# Split data into training and testing sets
split_index <- createDataPartition(Playoffs_combined$Series_ORtg, p = 0.8, list = FALSE)
train_data <- Playoffs_combined[split_index, ]
test_data <- Playoffs_combined[-split_index, ]

# Define target and features for training
x_train <- train_data %>% select(-Series_ORtg)
y_train <- train_data$Series_ORtg

# Train the random forest model
rf_model <- randomForest(x_train, y_train, importance = TRUE, ntree = 500)

# Ensure the test data contains the same columns as training
x_test <- test_data[, colnames(x_train), drop = FALSE]
y_test <- test_data$Series_ORtg

# Make predictions
predictions <- predict(rf_model, x_test)

# Calculate RMSE
rmse <- sqrt(mean((predictions - y_test)^2))
cat("Root Mean Squared Error:", rmse, "\n")

# Display variable importance
print(importance(rf_model))
varImpPlot(rf_model)
```
